---
title: "Team Project 1 - EDA"
author: "Maobin Guo"
output:
  pdf_document: default
  word_document: default
  html_document:
    
    df_print: paged
---
  
```{r start, echo = FALSE, include=FALSE }

options(warn=-1)
options(xtable.comment = FALSE)


library(ggplot2)
library(xtable)
library(rms)
library(arm)
library(pROC)
#library(e1071)
#library(caret)
#library(ggplot2)
#require(gridExtra)
library(dplyr)
library(reshape2)
library(ggalluvial)
```


```{r load,  echo = FALSE, include=FALSE}

###### Load the data
df <- read.csv("lalondedata.csv",header=T)

str(df)
summary(df)

#############################
# Type transform 

for (fac_var in c('black', 'hispan', 'married', 'nodegree', 'treat') ) {
  df[,fac_var] <- as.factor(as.character(df[,fac_var]))
}

df$job78 <- rep(0, nrow(df))
df$job78[df$re78 > 0 ] = 1  
df$job78 <- as.factor(df$job78)
levels(df$job78) <- c("unemployed", "employed")


df$job74 <- rep(0, nrow(df))
df$job74[df$re74 > 0 ] = 1  
df$job74 <- as.factor(df$job74)
levels(df$job74) <- c("unemployed", "employed")
```

## Variables 


#### 'black', 'hispan', 'married', 'nodegree'

```{r ,  echo = FALSE}

for (fac_var in c('black', 'hispan', 'married', 'nodegree', 'treat') ) {
  print(paste("Frequent of :", fac_var))
  
  print(table(df[fac_var]))
}

```

### Educ 

####  Educ overview
```{r ,  echo = FALSE}
hist(as.numeric(df$educ))
```



### Age 

####  Age overview
```{r ,  echo = FALSE}
hist(as.numeric(df$age))
```


####  Categorize age  
 
Categorize age to 3 group 

- [1] 15-24 : Those just entering the labour market following education
- [2] 25-54 : Those in their prime working lives
- [3] 55-64 : Passing the peak of their career and approaching retirement

Since the expeirment only take 2 years, year shifting was not consideding here. 

[Refere](https://data.oecd.org/emp/employment-rate-by-age-group.htm)

```{r ,  echo = FALSE}

df$age_fac <- rep(0, nrow(df))
df$age_fac[df$age >= 25 & df$age < 55] <- 1
df$age_fac[df$age >= 55] <- 2

df$age_fac <- factor(df$age_fac)
levels(df$age_fac) <- c("Rookie", "Prime", "Aged")

table(df$age_fac)
```

### Pivot views

#### Received treat conditions across differtn variable

```{r ,  echo = FALSE}
df %>% 
group_by(age_fac) %>% 
summarize(treatNum = sum(treat==1), total=n(), rate = round(100 * sum(treat==1)/n(),2))
```
**Insight**
Almost even in different working age group. 


```{r ,  echo = FALSE}
df %>% 
group_by(black) %>% 
summarize(treatNum = sum(treat==1), total=n(), rate = round(100 * sum(treat==1)/n(),2))
```
**Insight**
Black's treat rate is far more higher than unmarried. 



```{r ,  echo = FALSE}
df %>% 
group_by(hispan) %>% 
summarize(treatNum = sum(treat==1), total=n(), rate = round(100 * sum(treat==1)/n(),2))
```
**Insight**
Hispan's treat rate is lower than other race. 


```{r ,  echo = FALSE}
df %>% 
group_by(married) %>% 
summarize(treatNum = sum(treat==1),  total=n(), rate = round(100 * sum(treat==1)/n(),2))
```

**Insight**
Married people's treat rate is far more lower than unmarried. 

```{r ,  echo = FALSE}
df %>% 
group_by(nodegree) %>% 
summarize(treatNum = sum(treat==1),  total=n(), rate = round(100 * sum(treat==1)/n(),2))
```
**Insight**
No degree group treat rate is higher. 


```{r ,  echo = FALSE}
ggplot(data = df,
       aes(axis1 = job74, axis2 = treat, axis3 = job78)) +
  scale_x_discrete(limits = c("job74", "treat", "job78"), expand = c(.01, .05)) +
  geom_alluvium(aes(fill = job74)) +
  geom_stratum() + geom_text(stat = "stratum", aes(label = after_stat(stratum)))

```


```{r ,  echo = FALSE}

df %>%
group_by(job74, treat, job78) %>%
summarize(treatNum = n())

```

-  For Unemployed people in 74:
  - Treated    : 131  
    - Employed   in 78 : 100  (76.3%)
    - Unemployed in 78 : 31    (23%)
  - No treated : 112 
    - Employed   in 78 : 75    (67%)
    - Unemployed in 78 : 37    (33%)


-  For Employed people in 74:
  - Treated    : 54  
    - Employed   in 78 : 40    (74%)
    - Unemployed in 78 : 14    (26%)
  - No treated : 317 
    - Employed   in 78 : 256    (80%)
    - Unemployed in 78 : 61     (19.2%)


**Insight**

-  For umemployed people in 74,  treat seems improve employed rate (67% -> 76%)
-  For umemployed people in 74,  treat seems bring down employed rate (80% ->74%)
   Maybe the employed people in 74 and not treated people are people no need to take treat.
   
   
##### Only consider treat :

```{r ,  echo = FALSE}

df %>%
group_by(treat, job78) %>%
summarize(treatNum = n())

```

- For treated : 
  - unemployed :  45   (0.24)
  - employed :    140  (0.76)

- For untreated : 
  - unemployed :  98   (0.228)
  - employed :    331  (0.77)
  
**Insight**

If only take treated as a variable, it's hard to compare. 



   
```{r ,  echo = FALSE, out.width=c('25%', '25%', '25%', '25%'), fig.show='hold'}

df %>%
group_by(job74, treat, nodegree) %>%
summarize(treatNum = n())

ggplot(df, aes(x = re74, y = re78)) +
      geom_point(alpha = .5,colour="blue4") +
      geom_smooth(method="loess",col="red3")  

# exclude 2 outliers 
ggplot(df[ (df$re78 != max(df$re78)) & (df$re74 != max(df$re74)), ], 
      aes(x = re74, y = re78)) +
      geom_point(alpha = .5,colour="blue4") +
      geom_smooth(method="loess",col="red3")


ggplot(df, aes(x = educ, y = re78)) +
      geom_point(alpha = .5,colour="blue4") +
      geom_smooth(method="loess",col="red3")  


ggplot(df, aes(x = treat, y = re78,fill=re78)) +
  geom_boxplot() +
  labs(title="re78 & treat")

ggplot(df, aes(x = black, y = re78,fill=re78)) +
  geom_boxplot() +
  labs(title="re78 & black")


ggplot(df, aes(x = hispan, y = re78,fill=re78)) +
  geom_boxplot() +
  labs(title="re78 & hispan")


ggplot(df, aes(x = married, y = re78,fill=re78)) +
  geom_boxplot() +
  labs(title="re78 & married")

ggplot(df, aes(x = nodegree, y = re78,fill=re78)) +
  geom_boxplot() +
  labs(title="re78 & nodegree")


ggplot(df, aes(x = age_fac, y = re78,fill=re78)) +
  geom_boxplot() +
  labs(title="re78 & age_fac")

```


```{r ,  echo = FALSE}

ggplot(df, aes(x = treat, y = re78,fill=re78)) +
  geom_boxplot() +
  labs(title="re78 & treat ON black") +
  facet_wrap( ~ black,ncol=2)
```
** Insight **
- For balck people, treat is useful


```{r ,  echo = FALSE}

ggplot(df, aes(x = treat, y = re78, fill=re78)) +
  geom_boxplot() +
  labs(title="re78 & treat ON hispan") +
  facet_wrap( ~ hispan, ncol=2)
```
** Insight **
- No interaction
- Treat bring down re78



```{r ,  echo = FALSE}

ggplot(df, aes(x = treat, y = re78, fill=re78)) +
  geom_boxplot() +
  labs(title="re78 & treat ON married") +
  facet_wrap( ~ married, ncol=2)
```
** Insight **
- No interaction
 
```{r ,  echo = FALSE}

ggplot(df, aes(x = treat, y = re78, fill=re78)) +
  geom_boxplot() +
  labs(title="re78 & treat ON nodegree") +
  facet_wrap( ~ nodegree, ncol=2)
```
** Insight **
- No interaction
 

```{r ,  echo = FALSE}
ggplot(df, aes(x = age_fac, y = re78,fill=re78)) +
  geom_boxplot() +
  labs(title="re78 & age_fac") +
  facet_wrap( ~ nodegree, ncol=3)
```
** Insight **
- With interaction
- Treat will bring down the advantage of primary work. Maybe for low salary job, primay age's experience can be got from training. 



```{r ,  echo = FALSE, include=FALSE}

df$re74_sqr <- df$re74^2

model_raw <- lm(re78~treat+age_fac+educ+black+hispan+married+nodegree+re74 ,data=df)

model_raw2 <- lm(re78~treat+age_fac+educ+black+hispan+married+nodegree+re74+re74_sqr ,data=df)


#===============
model_raw3 <- lm(re78~treat * (black+re74+re74_sqr) + educ + hispan + married + nodegree +age_fac ,data=df)
summary(model_raw3)


model_new <- lm(re78~ treat * (re74 + re74_sqr) + black  + educ + hispan + married + nodegree + age_fac ,data=df)
summary(model_new)

anova(model_raw3, model_new)


#Auto 

model_null <- lm(re78~1, data=df)

model_aic_stepwise  <- step(model_raw3, scope = formula(model_null), direction="both", trace=0)
summary(model_aic_stepwise)

model_bic_stepwise <- step(model_null, scope=formula(model_raw3), direction="both", trace=0, k = log(nrow(df)))
summary(model_bic_stepwise)


model_forward_aic <- step(model_null, scope = formula(model_raw3),direction="forward", trace=0)
summary(model_forward_aic)

model_forward_bic <- step(model_null, scope = formula(model_raw3),direction="forward", trace=0,  k = log(nrow(df)))
summary(model_forward_bic)

model_back_aic <- step(model_null, scope = formula(model_raw2),direction="backward", trace=0)
summary(model_forward_aic)

model_back_bic <- step(model_null, scope = formula(model_raw2),direction="backward", trace=0,  k = log(nrow(df)))
summary(model_forward_bic)

```

## Recommand modle

```{r ,  echo = TRUE}

model_recomand <- lm(formula = re78 ~ treat + black + re74 + re74_sqr + educ + age_fac + treat:re74, data = df)
summary(model_recomand)

```

 


