---
title: "Assignment 3"
author: "Maobin Guo"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
    
---

```{r setup, echo = FALSE, include=FALSE}

options(warn=-1)
options(xtable.comment = FALSE)


library(ggplot2)
library(xtable)
library(rms)
library(arm)
library(pROC)
library(e1071)
library(caret)
library(ggplot2)
require(gridExtra)

```


```{r echo = FALSE, include=FALSE}
df <- read.csv("smoking.csv", stringsAsFactors = FALSE, sep = "," ,dec="," )

#Remove gestation
#df = subset(df, select = -c(date, gestation))

# collapse race categories from 0 - 5 into one category for race = white
df[df$mrace < 6, ]$mrace = 5
df[df$med <= 7 & df$med >= 6, ]$med = 7


# Factorize 
df$med   <- as.factor(df$med)
df$mrace <- as.factor(df$mrace)
df$smoke <- as.factor(df$smoke)
df$Premature_fac <- factor(as.integer(df$Premature), levels = c(0, 1), labels = c("pre", "normal"))
df$Premature <- as.numeric(df$Premature) 

df$parityCent <- df$parity - mean(df$parity)
df$mageCent <- df$mage - mean(df$mage)
df$mhtCent <- df$mht - mean(df$mht)
df$mpregwtCent <- df$mpregwt - mean(df$mpregwt)

```
 

```{r echo = FALSE}
 

```

\hfill\break

## Summary

In this investigation, a logistic regression model was built to quest the relationship between smoking and pre-term birth. Other factors would also be analyzed to check whether they could contribute to pre-term birth or change the relationship between smoking and premature delivery.

\hfill\break

## Introduction

This analysis based on a dataset from a study carried between 1960 and 1967 at the Kaiser Foundation Hospital in Oakland, CA. `r nrow(df)` observations were included in the analysis. Logistic regression was exploited as the primary technique to quest the answer to the following questions. 

- Is there any relationship between pre-term birth and mother's smoking? If any, how much the difference of odds ratio of pre-term birth between smoking and nonsmoking mother. 

- Would the mother's race have a significant influence on the correlation?

- Apart from smoking, are there any other factors in the data that would contribute to pre-term birth. 

\hfill\break

## Data

\hfill\break

### Interaction Explore 

```{r echo = FALSE, include=FALSE} 
# Hiddend analysis 

ggplot(df,aes(x=Premature_fac, y=parity, fill=Premature_fac)) +
  geom_boxplot() + coord_flip() +
  scale_fill_brewer(palette="Reds") +
  labs(title="Mother's parity vs Premature",
       x="Premature",y="Mother's parity") + 
  theme_classic() + theme(legend.position="none")

ggplot(df,aes(x=Premature_fac, y=mage, fill=Premature_fac)) +
  geom_boxplot() + coord_flip() +
  scale_fill_brewer(palette="Reds") +
  labs(title="Mother's age vs Premature",
       x="Premature",y="Mother's age") + 
  theme_classic() + theme(legend.position="none")

ggplot(df,aes(x=Premature_fac, y=mht, fill=Premature_fac)) +
  geom_boxplot() + coord_flip() +
  scale_fill_brewer(palette="Reds") +
  labs(title="Mother's height vs Premature",
       x="Premature",y="Mother's height") + 
  theme_classic() + theme(legend.position="none")

ggplot(df,aes(x=Premature_fac, y=inc, fill=Premature_fac)) +
  geom_boxplot() + coord_flip() +
  scale_fill_brewer(palette="Reds") +
  labs(title="Mother's income vs Premature",
       x="Premature",y="Mother's income") + 
  theme_classic() + theme(legend.position="none")

```


```{r echo = FALSE, out.width=c('25%', '25%', '25%', '25%'), fig.show='hold'}

#------------------------------------

ggplot(df,aes(x=Premature_fac, y=parity, fill=Premature_fac)) +
  geom_boxplot() + coord_flip() +
  scale_fill_brewer(palette="Reds") +
  labs(title="Mother's parity vs Premature",
       x="Premature",y="Mother's parity") + 
  theme_classic() + theme(legend.position="none") +
  facet_wrap( ~ smoke)


ggplot(df,aes(x=Premature_fac, y=mage, fill=Premature_fac)) +
  geom_boxplot() + coord_flip() +
  scale_fill_brewer(palette="Reds") +
  labs(title="Mother's age vs Premature",
       x="Premature",y="Mother's age") + 
  theme_classic() + theme(legend.position="none") + 
  facet_wrap( ~ smoke)

ggplot(df,aes(x=Premature_fac, y=mht, fill=Premature_fac)) +
  geom_boxplot() + coord_flip() +
  scale_fill_brewer(palette="Reds") +
  labs(title="Mother's height vs Premature",
       x="Premature",y="Mother's height") + 
  theme_classic() + theme(legend.position="none") + 
  facet_wrap( ~ smoke)

ggplot(df,aes(x=Premature_fac, y=inc, fill=Premature_fac)) +
  geom_boxplot() + coord_flip() +
  scale_fill_brewer(palette="Reds") +
  labs(title="Mother's height vs Premature",
       x="Premature",y="Mother's income") + 
  theme_classic() + theme(legend.position="none") + 
  facet_wrap( ~ smoke)

```

- For the above four factors, their behavior on facilitating pre-term birth seems different with smoking and nonsmoking. Further investigation would be implemented in the modeling step. 


```{r echo = FALSE, include=FALSE}
# Hide 

# Explore Premature & factor variables

chisq.test(table(df[,c("Premature_fac","mrace")]))


apply(table(df[,c("Premature_fac","smoke")])/sum(table(df[,c("Premature_fac","smoke")])),  2,function(x) x/sum(x)) 

tapply(df$Premature_fac, df$smoke, function(x) table(x)/sum(table(x)))
chisq.test(table(df[,c("Premature_fac","smoke")]))

chisq.test(table(df[,c("Premature_fac","med")]))

df$inc_fac <- as.factor(df$inc)
chisq.test(table(df[,c("Premature_fac","inc_fac")]))

```


###  Race  

Mother's race and education status have strong association with premature. 

```{r echo = FALSE, results='asis'}

xtable(apply(table(df[,c("Premature_fac","med")])/sum(table(df[,c("Premature_fac","med")])), 2,function(x) x/sum(x)) )

chi <- chisq.test(table(df[,c("Premature_fac","med")]))
```

- The chisq-test's p-value of mother's education vs prematrue is `r round(chi$p.value, 4)`, which mean there is a strong tie between mother's educational experience and premature rate 

- There is some noticeable difference in the premature rate between mothers with diverse educational backgrounds.  Mothers with college or some college experience are more likely premature than others. 

### Education

```{r echo = FALSE, results='asis'}

xtable(apply(table(df[,c("Premature_fac","mrace")])/sum(table(df[,c("Premature_fac","mrace")])), 2,function(x) x/sum(x)) )

chi <- chisq.test(table(df[,c("Premature_fac","mrace")]))
```

- The chisq-test's p-value of mother's ramce vs prematrue is `r round(chi$p.value, 4)`, which mean there is a strong associaton between mother's race and premature rate.  

- Comparing with other rame, mix race has a much higher probability of having pre-term birth. 

- Asian mother's premature rate is remarkably lower than mothers of other races. 


###  Smoking

```{r echo = FALSE, results='asis'}

xtable(apply(table(df[,c("Premature_fac","smoke")])/sum(table(df[,c("Premature_fac","smoke")])), 2,function(x) x/sum(x)) )

chi <- chisq.test(table(df[,c("Premature_fac","smoke")]))

```

- The chisq-test's p-value of mother's smoking vs. premature is `r round(chi$p.value, 4)`, which mean the association is extremely weak or even not exist

- This is a somewhat unexpected finding; typically, the common sense is that smoking would bring various disadvantages to health. To be confirmed, the result would be investigated in the following steps. 

## Model


```{r echo = FALSE , include=FALSE}

raw_model <- glm(Premature_fac ~ parityCent + mageCent + mhtCent + mpregwtCent + smoke + inc + mrace + med , data = df, family = binomial)


null_model <- glm(Premature_fac ~1 , data = df, family = binomial)


full_model <- glm(Premature_fac ~ parityCent + mageCent + mhtCent + mpregwtCent + smoke + inc + mrace + med + smoke:med + smoke +mrace, data = df, family = binomial)


model  <- step(full_model, scope = formula(null_model), direction="forward",trace=0)

summary(model)
 
 
res1 <- residuals(full_model,"resp")

binnedplot(x=fitted(model),y=res1,xlab="Pred. probabilities",
           col.int="red4",ylab="Avg. residuals",main="Binned residual plot",col.pts="navy")

binnedplot(x=df$parityCent,y=res1,xlab="Arsenic centered",
           col.int="red4",ylab="Avg. residuals",main="Binned residual plot",col.pts="navy")

binnedplot(x=df$mageCent,y=res1,xlab="Arsenic centered",
           col.int="red4",ylab="Avg. residuals",main="Binned residual plot",col.pts="navy")

binnedplot(x=df$mhtCent,y=res1,xlab="Arsenic centered",
           col.int="red4",ylab="Avg. residuals",main="Binned residual plot",col.pts="navy")

binnedplot(x=df$mpregwtCent,y=res1,xlab="Arsenic centered",
           col.int="red4",ylab="Avg. residuals",main="Binned residual plot",col.pts="navy")

```

```{r echo = FALSE , include=FALSE}

full_model <- glm(Premature_fac ~ parityCent + mageCent + mhtCent + mpregwtCent + smoke + inc  + med  + mrace, data = df, family = binomial)

res1 <- residuals(full_model,"resp")

model_sum <- summary(full_model)

```

The  modle is :

$$
logit(\pi_{i}) = \beta_{0} + \beta_{1}*parityC + \beta_{2}*mageC + \beta_{3}*mhtC +\beta_{4}*mpregwtC + \beta_{5}*smoke +\beta_{6}*inc +\beta_{7}*med +\beta_{8}*mrace
$$

- parityC :  cented parity
- mageC   :  cented mage
- mhtC    :  cented mht
- mpregwtC : cented mpregwtC 


### Model performance

```{r echo = FALSE, results='asis'}

xtable(model_sum)

```

Null deviance : `r round(model_sum$null.deviance,2)`

Residual deviance : `r round(model_sum$deviance,2)`



### Model diagnostics


```{r echo = FALSE, out.width=c('25%', '25%', '25%', '25%'), fig.show='hold'}

binnedplot(x=fitted(model),y=res1,xlab="Pred. probabilities",
           col.int="red4",ylab="Avg. residuals",main="Binned residual plot",col.pts="navy")

binnedplot(x=df$mageCent,y=res1,xlab="Arsenic centered",
           col.int="red4",ylab="Avg. residuals",main="Binned residual plot",col.pts="navy")

binnedplot(x=df$mhtCent,y=res1,xlab="Arsenic centered",
           col.int="red4",ylab="Avg. residuals",main="Binned residual plot",col.pts="navy")

binnedplot(x=df$mpregwtCent,y=res1,xlab="Arsenic centered",
           col.int="red4",ylab="Avg. residuals",main="Binned residual plot",col.pts="navy")

```

### Model validation
 
###### ROC 
 

```{r echo = FALSE, out.height= c('30%'), message=FALSE}

roc <- roc(df$Premature,fitted(full_model),plot=T,print.thres="best",legacy.axes=T, col="blue2")

roc_value <-coords(roc, "best", transpose = FALSE)

```



**AUC :** `r round(roc$auc, 2)`

**Cut-off :** `r round(roc_value[1], 2)`


```{r echo = FALSE, results='asis', include=FALSE}

Conf_mat <- confusionMatrix(as.factor(ifelse(fitted(full_model) >= as.numeric(roc_value[1]), "1","0")), as.factor(df$Premature),positive = "1")

```


\hfill\break

##### Confusion Matrix 

\hfill\break

```{r echo = FALSE, results='asis'}
xtable(Conf_mat$table)
```

\hfill\break

**Sensitivity :** `r round(Conf_mat$byClass[1], 2)`

**Specificity :** `r round(Conf_mat$byClass[2], 2)`


\hfill\break

## Conclusion












