---
title: "Assignment 5"
author: "Maobin Guo"
output:
  pdf_document: default
  word_document: default
  html_document:
    
    df_print: paged
---

```{r setup, echo = FALSE, include=FALSE}

options(warn=-1)
options(digits=2)
options(xtable.comment = FALSE)

library(ggplot2)
library(xtable)
library(sjPlot)
library(car)
library(cobalt)
library(MatchIt)
library(randomForest)

```

# Part I

```{r echo = FALSE, include=FALSE}
df <- read.csv("./data/Asthma.txt", 
               sep = " ",
               stringsAsFactors = FALSE, 
                strip.white=TRUE)

df[df$pg==2, ]$pg = 0


df$pg <- as.factor(df$pg)
df$i_race <- as.factor(df$i_race)
df$i_insu <- as.factor(df$i_insu)
df$i_drug <- as.factor(df$i_drug)
df$i_aqoc <- as.factor(df$i_aqoc)

df$com_t  <- df$com_t  - mean(df$com_t)
df$pcs_sd <- df$pcs_sd - mean(df$pcs_sd)
df$mcs_sd <- df$mcs_sd - mean(df$mcs_sd)

df$i_sex  <- relevel(factor(df$i_sex),  ref = 1)
df$i_educ <- relevel(factor(df$i_educ), ref = 5)
df$i_seve <- relevel(factor(df$i_seve), ref = 3)

summary(df)
str(df)

#miss_data <- df




```
 
## Answer 1

```{r echo = FALSE, include=FALSE}

# pg (treatment assignment)
# i_aqoc  outcome

asd <- bal.tab(list(treat=df$pg,covs=df[,2:10],estimand="ATT"))
asd$Balance[abs(asd$Balance$Diff.Un) > 0.1, ] 
```

The following covariates of this dataset are not balanced because the absolute value of Diff.Un for this variables are bigger than 0.1

- i_sex
- i_race
- i_educ
- com_t
- pcs_sd

```{r echo = FALSE, results='asis', message=FALSE , warning=FALSE}
xtable(asd$Balance[abs(asd$Balance$Diff.Un) > 0.1, ])
```



```{r echo = FALSE, include=FALSE, results='asis', message=FALSE}

m1 <- glm(pg ~ i_age + i_sex + i_race + i_educ +
               i_insu + i_drug + i_seve + com_t + 
               pcs_sd + mcs_sd, 
          data = df,
          family = binomial)
summary(m1)

pscores <- predict(m1, type = "response")
summary(pscores)

ggplot(df, 
       aes(pscores)) +
       geom_histogram(alpha=.6,fill=rainbow(10),bins=10)

df$pscores <- pscores

```


## Answer 2.a

```{r echo = FALSE, include=FALSE, results='asis', message=FALSE}

#look at distribution of propensity scores for treateds and controls
 
#ggplot(df, aes(y=pscores, x=pg, fill=pg)) + geom_boxplot()
#ggplot(df, aes(x=pscores, fill=pg)) + geom_density(alpha=.3) + xlim(0, 1)

range_p2 <- c(min(df[df$pg == 0, ]$pscores), max(df[df$pg == 0, ]$pscores))

range_p1 <- c(min(df[df$pg == 1, ]$pscores), max(df[df$pg == 1, ]$pscores))

# P2 whoes propensitiy score is not in P1's propensitiy score range

df[(df$pg == 0) & ((df$pscores < range_p1[1]) | (df$pscores > range_p1[2] )), ]
df[(df$pg == 1) & ((df$pscores < range_p2[1]) | (df$pscores > range_p2[2] )), ]


dfol <- rbind(
 df[(df$pg == 0) & ((df$pscores > range_p1[1]) & (df$pscores < range_p1[2] )), ], 
 df[(df$pg == 1) & ((df$pscores > range_p2[1]) & (df$pscores < range_p2[2] )), ]
)


```

There are `r dim(df)[1] - dim(dfol)[1]` observations with an estimated propensity score that is out of the range of e in the other group. Accroding to instruction, these
observations were discarded. 


## Answer 2.b

Compare with raw data, the balance of matched data is much better. The following
covariance becomes balanced:

- i_sex
- i_race

The diff value of i_educ still slightly bigger than 0.1, however it is much 
balance than its diff value in raw data. 


```{r echo = FALSE, include=FALSE}
###### Propensity scores matching
#main call-- embed the logistic regression inside the call
#start with a main effects only regression
matches_res <- matchit(pg ~ i_age + i_sex + i_race + i_educ +
                          i_insu + i_drug + i_seve + com_t + 
                          pcs_sd + mcs_sd,
                          method = "nearest", 
                          distance = "logit", 
                          data = dfol)

#df$pscores = pscores
#ggplot(matched_df, aes(y=distance, x=pg, fill=pg)) +
#  geom_boxplot()

matched_df  <- match.data(matches_res)

matched_asd <- bal.tab(list(treat=matched_df$pg, 
                            covs=matched_df[,2:10],
                            estimand="ATT"))

matched_asd$Balance[abs(matched_asd$Balance$Diff.Un) > 0.1, ]
 
```

```{r echo = FALSE, include=TRUE, results='asis', message=FALSE, warning=FALSE}
xtable(matched_asd$Balance[abs(matched_asd$Balance$Diff.Un) > 0.1, ])
```
 


## Answer 2.c

```{r echo = FALSE, include=FALSE, results='asis', message=FALSE}

#p1 <- predict(m1, type = "response", newdata = matched_df[matched_df$pg==1, ])
#p2 <- predict(m1, type = "response", newdata = matched_df[matched_df$pg==0, ])

#se <- sqrt(var(minwagematcheddata$EmploymentPost[minwagematcheddata$PA.NJ==1])/73 + 
#            var(minwagematcheddata$EmploymentPost[minwagematcheddata$PA.NJ==0])/73)

n = nrow(matched_df)/2

diff_mean <- mean((as.numeric(matched_df$i_aqoc[matched_df$pg==0]) - 1)/n) - 
             mean((as.numeric(matched_df$i_aqoc[matched_df$pg==1]) - 1)/n)


se <- sqrt(var((as.numeric(matched_df$i_aqoc[matched_df$pg==0]) - 1)/n) + 
           var((as.numeric(matched_df$i_aqoc[matched_df$pg==1]) - 1)/n))

diff_ci = c(diff_mean - 1.96 * se, diff_mean + 1.96 * se)

```

Base on the 1-on-1 data, the SE is `r se ` and the 95% confidence interval is  
[`r diff_ci`]. Because the confidence interval contains 0, we cannot claim that 
different physician group cause the service difference.


## Answer 2.d

?? use which dataset?  matched or whole ?
?? Is pre-treatment include pysical assignment ?
?? Need AIC steps search ?

```{r echo = FALSE, include=FALSE, results='asis', message=FALSE}

m5 <- glm(i_aqoc ~ i_age + i_sex + i_race + i_educ +
                   i_insu + i_drug + i_seve + com_t + 
                   pcs_sd + mcs_sd + pscores, 
          data = dfol,
          family = binomial)
summary(m5)


m6 <- glm(i_aqoc ~ i_age + i_sex + i_race + i_educ +
                   i_insu + i_drug + i_seve + com_t + 
                   pcs_sd + mcs_sd + pscores, 
          data = matched_df,
          family = binomial)
summary(m6)




```

## Answer 2.e
?? whay control sample size is bigger than another gourp ?

Sample sizes:
          Control Treated
All           133      97
Matched       102      97
Unmatched      31       0
Discarded       0       0


```{r echo = FALSE, include=FALSE, results='asis', message=FALSE}

matches_res <- matchit(pg ~ i_age + i_sex + i_race + i_educ +
                          i_insu + i_drug + i_seve + com_t + 
                          pcs_sd + mcs_sd,
                          method = "nearest", 
                          distance = "logit", 
                          replace=TRUE,
                          ratio =5,
                          data = dfol)
 

matched_df  <- match.data(matches_res)

matched_asd <- bal.tab(list(treat=matched_df$pg, 
                            covs=matched_df[,2:10],
                            estimand="ATT"))

matched_asd$Balance[abs(matched_asd$Balance$Diff.Un) > 0.1, ]


n = nrow(matched_df)/2

diff_mean <- mean((as.numeric(matched_df$i_aqoc[matched_df$pg==0]) - 1)/n) - 
             mean((as.numeric(matched_df$i_aqoc[matched_df$pg==1]) - 1)/n)


se <- sqrt(var((as.numeric(matched_df$i_aqoc[matched_df$pg==0]) - 1)/n) + 
           var((as.numeric(matched_df$i_aqoc[matched_df$pg==1]) - 1)/n))

diff_ci = c(diff_mean - 1.96 * se, diff_mean + 1.96 * se)

```

```{r echo = FALSE, include=TRUE, results='asis', message=FALSE, warning=FALSE}
xtable(matched_asd$Balance[abs(matched_asd$Balance$Diff.Un) > 0.1, ])


```

## Answer 3

The method  one-to-one is better. Why ? ....




